Stochastic Variance-Reduced Policy Gradient

### 2018-02-08 svrpg.tex
reverted last change
Minor fixes

\begin{abstract}
        In this paper, we propose a stochastic variance-reduced version of policy gradient for solving Markov Decision Processes (MDPs).
        Stochastic variance-reduced gradient (SVRG) methods have proven to be very successful in supervised learning.
        However, their adaptation to policy gradient is not straightforward and needs to account for I) a non-concave objective function; II) approximations in the full gradient computation; and III) a non-stationary sampling process.
        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
        We further analyse its convergence guarantees under standard assumptions on the MDP.
        Finally, we suggest practical variants of SVRPG, and we empirically evaluate them on continuous MDPs.
\end{abstract}

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{see Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. {\color{red}In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with non-concave objective.} 
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

@@

We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}

@@

The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties:

@@

Note that it is important that the trajectories used in the second and the third term are the same for the variance to vanish.

@@

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG with REINFORCE or G(PO)MDP gradient estimator.
We mainly leverage on the recent analysis of non-concave SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In Appendix~\ref{app:gauss} we show how the traditional Gaussian policy satisfies these assumptions.

@@

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Assume the REINFORCE or the G(PO)MDP gradient estimator is used in SVRPG (see Equation~\eqref{E:svrpg.estimate.batch}).
Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B},
\end{align*}
where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to Appendix~\ref{app:proofs} for a detailed proof involving the definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$. In this sense, advantage over batch gradient ascent can be achieved with properly selected meta-parameters. In Section~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
Finally, from the return statement of Algorithm \ref{alg:svrpg}, it is worth noting that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm. This is particularly meaningful in the context of online learning that we are considering in this paper.

@@

\subsection{Zeroth update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data. In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following we will always assume this practical variant is employed.

\subsection{Meta-Parameter Selection}\label{sec:stopping}
% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
The step size $\alpha$ is crucial to balance variance reduction and efficiency.
% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
Epoch length $m$ influences the variance introduced by the importance weights: low values of $m$ are associated with low variance but increase the frequency of snapshot points (which means many FG computations); high values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing high variance in the importance weights. We propose to jointly adapt the two meta-parameters.
% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.

\textbf{Adaptive step size.}
A standard way to deal with noisy gradients is to use adaptive strategies to compute the step size.
ADAptive Moment estimation (ADAM)~\citep{kingma2014adam} stabilizes the parameter update by computing learning rates for each parameter based on an incremental estimate of the gradient variance.
Due to this feature, we would like to incorporate ADAM in the structure of the SVRPG update.
Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
Given this structure, we suggest to use two separate ADAM estimators:
\begin{align*}
        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} v_t^{s+1} \text{ for $t=1,\dots,m-1$},
\end{align*}
where $\alpha^{\textsc{FG}}_{s}$ is associated to the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
The use of two ADAM estimators allows to capture and exploit this property.

\textbf{Adaptive epoch length.}
It is easy to imagine that a predefined schedule (\eg $m$ fixed in advance or changed with a policy-independent process) may poorly perform due to the high variability of the updates.
In particular, given a fixed number of sub-iterations $m$, the variance of the updates in the sub-iterations depends on the snapshot policy and the sampled trajectories.
Since the ADAM estimate partly captures such variability,  we propose to take a new snapshot (\ie interrupt the sub-iterations) whenever the step size $\alpha^{\textsc{SI}}$ proposed by ADAM for the sub-iterations is smaller than the one for the FG (\ie $\alpha^{\textsc{FG}}$).
If the latter condition is verified, it amounts to say that the noise in the corrected gradient has overcome the information of the FG.
Formally, the stopping condition is as follow
\[
        \textbf{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot,}
\]
where we have introduced $N$ and $B$ to take into account the trajectory efficiency (\ie weighted advantage).
The less the number of trajectories used to update the policy, the better.
This weighted approach comes from the fact that, as mentioned above, there is no advantage in having $m$ too small.
Thus, we aim to facilitate sub-iterations rather than snapshots.

\subsection{Normalized Importance Sampling}\label{sec:prac}
As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization \citep{precup2000eligibility}.
This technique can reduce the variance of the importance weights at the cost of introducing some bias. Whether the trade-off is advantageous depends on the specific task.  
Introducing self-normalization in the context of our algorithm, we switch from Eq.~\eqref{E:svrpg.estimate.batch} to:
\begin{align*}
\blacktriangledown J(\vtheta_{t}) = \wh{\nabla}_N J(\wt{\vtheta})
 + \frac{1}{B} \sum_{i=0}^{B-1} \left[
g(\tau_i|\vtheta_t)\right]\\ 
- \frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right].
\end{align*}
In Section \ref{sec:exp} we show some cases in which self-normalization provides a performance improvement.

@@

\section{Experiments}\label{sec:exp}
In this section we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks. This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}). Each experiment is run 10 times with a random policy initialization, different each time, shared by all the algorithms under comparison. The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. Of course test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.

>>>>

\begin{abstract}
        In this paper, we propose a novel reinforcement-learning algorithm consisting in a stochastic variance-reduced version of policy gradient for solving Markov Decision Processes (MDPs).
        Stochastic variance-reduced gradient (SVRG) methods have proven to be very successful in supervised learning.
        However, their adaptation to policy gradient is not straightforward and needs to account for I) a non-concave objective function; II) approximations in the full gradient computation; and III) a non-stationary sampling process.
        The result is SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
        We further analyse its convergence guarantees under standard assumptions on the MDP.
        Finally, we suggest practical variants of SVRPG, and we empirically evaluate them on continuous MDPs.
\end{abstract}

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems, the usual approach is to tune the batch size of SG to find the optimal trade-off between variance and speed.
Recall that, compared to SL, the samples are not fixed in advance but we need to collect them at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
For these reasons, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following issues:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The objective function is an expected value over the trajectory density $p_{\vtheta}(\tau)$ of the total discounted reward, for which we would need an infinite dataset.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the policy parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as Gaussian\footnote{see Appendix~\ref{app:gauss} for more details on the Gaussian policy case.} and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being concave, showing that SVRG is robust to an inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. In Section \ref{sec:conv} we will show how the estimation accuracy impacts on the convergence results with a non-concave objective.
Finally, the non-stationarity of the optimization problem introduces a bias into the SVRG estimator in Eq.~\eqref{E:svrg.gradient.correction}.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

@@

It is worth noting that the full gradient and the correction term have the same expected value: $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{A:gradient_estimators} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}

@@

The use of mini-batches is also common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties:

@@

Note that it is essential that the trajectories used in the second and the third term are the same for the variance to vanish.

@@

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG with REINFORCE or G(PO)MDP gradient estimator.
We mainly leverage on the recent analysis of non-concave SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Section~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In Appendix~\ref{app:gauss} we show how Gaussian policies satisfy these assumptions.

@@

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Assume the REINFORCE or the G(PO)MDP gradient estimator is used in SVRPG (see Equation~\eqref{E:svrpg.estimate.batch}).
Under Assumptions~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Algorithm~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B},
\end{align*}
where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to Appendix~\ref{app:proofs} for a detailed proof involving the definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
By analysing the upper-bound in Theorem~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on non-concave SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$. In this sense, an advantage over batch gradient ascent can be achieved with properly selected meta-parameters. In Section~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
Finally, from the return statement of Algorithm \ref{alg:svrpg}, it is worth noting that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm. This is particularly meaningful in the context of online learning that we are considering in this paper.

@@

\subsection{Zeroth update}
As noted in Section \ref{sec:alg}, the update performed at the beginning of each epoch is equivalent to a full-gradient update. In our setting, where collecting samples is particularly expensive, the $B$ trajectories collected using the snapshot trajectory $\pi_{\tilde{\vtheta}^s}$ feel like a waste of data. In practice, we just perform an approximate full gradient update using the $N$ trajectories sampled to compute $\gradApp{\tilde{\vtheta}^s}{N}$, \ie
\begin{align*}
	&\vtheta_{1}^{s+1} = \tilde{\vtheta}^s + \alpha\gradApp{\tilde{\vtheta}^s}{N} \\
	&\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v_t^{s+1} \text{ for $t=1,\dots,m-1$}.
\end{align*}
In the following, we will always assume this practical variant is employed.

\subsection{Meta-Parameter Selection}\label{sec:stopping}
% In section \ref{sec:conv}, we introduced the practical need to properly choose $\alpha$ and $m$, which are, respectively, the step size and the number of sub-iterations to perform after a snapshot.
% At this point, the open question is how to properly select the step size $\alpha$ and the sub-iteration number $m$ in practical applications.
The step size $\alpha$ is crucial to balance variance reduction and efficiency.
% The first parameter is crucial to balance the variance introduced by estimations within sub-iterations with respect to the variance introduced by estimations associated to the snapshots.
Epoch length $m$ influences the variance introduced by the importance weights. Low values of $m$ are associated with small variance but increase the frequency of snapshot points (which means many FG computations). High values of $m$ may move policy $\pi_{\vtheta_t}$ far away from the snapshot policy $\pi_{\wt{\vtheta}}$, causing large variance in the importance weights. We propose to adapt the two meta-parameters jointly.
% the farther the distance between the snapshot policy and the current one, the higher the variance of the importance weights.

\textbf{Adaptive step size.}
A standard way to deal with noisy gradients is to use adaptive strategies to compute the step size.
ADAptive Moment estimation (ADAM)~\citep{kingma2014adam} stabilizes the parameter update by computing learning rates for each parameter based on an incremental estimate of the gradient variance.
Due to this feature, we would like to incorporate ADAM in the structure of the SVRPG update.
Recall that SVRPG performs two different updates of the parameters $\vtheta$: I) FG update in the snapshot; II) corrected gradient update in the sub-iterations.
Given this structure, we suggest using two separate ADAM estimators:
\begin{align*}
        \vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\wh{\nabla}_N J(\wt{\vtheta}^s)\\
        \vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} v_t^{s+1} \text{ for $t=1,\dots,m-1$},
\end{align*}
where $\alpha^{\textsc{FG}}_{s}$ is associated with the snapshot and $\alpha^{\textsc{SI}}_{s+1,t}$ with the sub-iterations (see Appendix~\ref{app:practicalsvrpg} for details).
By doing so, we decouple the contribution of the variance due to the approximate FG from the one introduced by the sub-iterations.
Note that these two terms have different orders of magnitude since are estimated with a different number of trajectories ($B \ll N$) and w/o importance weights.
The use of two ADAM estimators allows to capture and exploit this property.

\textbf{Adaptive epoch length.}
It is easy to imagine that a predefined schedule (\eg $m$ fixed in advance or changed with a policy-independent process) may poorly perform due to the high variability of the updates.
In particular, given a fixed number of sub-iterations $m$, the variance of the updates in the sub-iterations depends on the snapshot policy and the sampled trajectories.
Since the ADAM estimate partly captures such variability,  we propose to take a new snapshot (\ie interrupt the sub-iterations) whenever the step size $\alpha^{\textsc{SI}}$ proposed by ADAM for the sub-iterations is smaller than the one for the FG (\ie $\alpha^{\textsc{FG}}$).
If the latter condition is verified, it amounts to say that the noise in the corrected gradient has overcome the information of the FG.
Formally, the stopping condition is as follows
\[
        \textbf{If }        \frac{\alpha^{\textsc{FG}}}{N} > \frac{\alpha^{\textsc{SI}}}{B} \textbf{ then } \text{take snapshot,}
\]
where we have introduced $N$ and $B$ to take into account the trajectory efficiency (\ie weighted advantage).
The less the number of trajectories used to update the policy, the better.
This weighted approach comes from the fact that, as mentioned above, there is no advantage in having $m$ too small.
Thus, we aim to facilitate sub-iterations rather than snapshots.

\subsection{Normalized Importance Sampling}\label{sec:prac}
As mentioned in Section~\ref{sec:stopping}, importance weights are an additional source of variance. A standard way to cope with this issue is self-normalization \citep{precup2000eligibility}.
This technique can reduce the variance of the importance weights at the cost of introducing some bias. Whether the trade-off is advantageous depends on the specific task.  
Introducing self-normalization in the context of our algorithm, we switch from Eq.~\eqref{E:svrpg.estimate.batch} to:
\begin{align*}
\blacktriangledown J(\vtheta_{t}) = \wh{\nabla}_N J(\wt{\vtheta})
 + \frac{1}{B} \sum_{i=0}^{B-1} \left[
g(\tau_i|\vtheta_t)\right]\\ 
- \frac{1}{ \sum_{i=0}^{B-1}\omega(\tau_i|\vtheta_t, \wt{\vtheta})} \sum_{i=0}^{B-1} \left[ \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right].
\end{align*}
In Section \ref{sec:exp} we show that self-normalization can provide a performance improvement.

@@

\section{Experiments}\label{sec:exp}
In this section, we evaluate the performance of SVRPG and compare it with policy gradient (PG) on well known continuous RL tasks. This comparison is performed using G(PO)MDP as an estimator of $\score{\vtheta}{\tau}\Reward(\tau)$ in both cases. For our algorithm, we use a batch size $N=100$, a mini-batch size $B=10$, and the jointly adaptive step size $\alpha$ and epoch length $m$ proposed in Section \ref{sec:stopping}. Since the aim of this comparison is to show the improvement that SVRG-flavored variance reduction brings to SG in the policy gradient framework, we set the batch size of the baseline policy gradient algorithm to $B$. In this sense, we measure the improvement yielded by computing snapshot gradients and using them to adjust parameter updates. Since we evaluate on-line performance over the number of sampled trajectories, the cost of computing such snapshot gradients is automatically taken into consideration. To make the comparison fair, we also use Adam in the baseline PG algorithm, which we will denote simply as G(PO)MDP in the following.
In all the experiments we used deep Gaussian policies with adaptive standard deviation (details on network architecture in Appendix \ref{app:exp}). Each experiment is run 10 times with a random policy initialization, different each time, shared by all the algorithms under comparison. The length of the experiment, \ie the total number of trajectories, is fixed for each task. Performance is evaluated by using test-trajectories on a subset of the policies considered during the learning process. Of course, test trajectories are not counted, since they are only meant to improve the accuracy of reported results and do not affect learning. We provide average performance with 90\% bootstrap confidence intervals.  
Task implementations are from the \textit{rllab} library \cite{duan2016benchmarking}\footnote{https://github.com/rll/rllab}, on which our agents are also based.
